---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# docker

<!-- badges: start -->
<!-- badges: end -->

This repo provides docker images on which to run code in *Geocomputation with R*.

The base image is `rockerdev/geospatial:3.6.3` from [github.com/rocker-org/rocker-versioned2](https://github.com/rocker-org/rocker-versioned2/blob/master/dockerfiles/Dockerfile_geospatial_3.6.3).

To build on different system configurations we provide tags that correspond to the following categories:

`baseimage-ubuntugis-setup-rpackages-buildbook`

The default base image is `rockerdev/geospatial:3.6.3`.
More images may be added in the future.


```{r}
baseimage = c(
  rockerdev_geospatial_3.6.3 = ""
)
```

Ubuntugis options include using the `ubuntugis-unstable` and `ubuntugis-stable` repos.

```{r}
ubuntugis = c(
  no_ubuntugis = "default_repos",
  ubuntugis_unstable = "ubuntugis_unstable",
  ubuntugis_stable = "ubuntugis_stable"
)
```

Setup options can include RStudio settings (yet to be added).

R package options relate to which R packages are installed on the image (yet to be added).

Buildbook options report whether or not the book is built:

```{r}
buildbook = c(
  no = "",
  yes = "buildbook"
)
```

We will create a 'build matrix' covering all combinations of these options (excluding the base image for now):

```{r}
g = expand.grid(ubuntugis, buildbook, stringsAsFactors = FALSE)
g
```

These can be converted into tags as follows:

```{r}
tag_df = tidyr::unite(g, tag)
tags = gsub(pattern = "__|^_|_$", replacement = "", tag_df$tag)
tags
```

We could write code to auto-generate Dockerfiles, as demonstrated in [rocker-org/rocker-versioned2](https://github.com/rocker-org/rocker-versioned2).

For now, to start the project going, we will manually edit the files, which can be created as follows:

```{r}
new_dockerfiles = paste0("Dockerfile_", tags)
lapply(new_dockerfiles, file.copy, from = "rockerdev-ubuntugis-bookbuild/Dockerfile", TRUE)
```

Edit these files as appropriate:

```{r, eval=FALSE}
file.edit("Dockerfile_ubuntugis_unstable")
```

